---
title: "Methods 2 -- Portfolio Assignment 1"
output:
  pdf_document: default
---

- _Type:_ Group assignment
- _Due:_ 10 March 2024, 23:59

---
```{r}
library(pacman)
p_load(ggplot2)
```

In the following exercises, you will be asked to generate and summarize simulations from statistical models. You should use what you have learned so far (i.e. for loop, if else statements, sampling from continuous and discrete distributions...) to generate observations and summarize your samples using (one of) the appropriate methods. You can find examples of how to do that in Ch. 5. Note that here we will only focus on generative models, several aspects for inference and hypothesis testing discussed in Ch. 4 are not directly needed in this context.

In the first exercise, we will assume that the population of interest has a proportion of 0.51 men and 0.49 women. Your model should reflect that.

Please submit your answers on GitHub Classroom.

---

1. _(5.2 from ROS)_ __Continuous probability simulation:__ The logarithms of weights (in pounds) of men in the United States are approximately normally distributed with mean 5.13 and standard deviation 0.17; women’s log weights are approximately normally distributed with mean 4.96 and standard deviation 0.20. Suppose 10 adults selected at random step on an elevator with a capacity of 1750 pounds. What is the probability that their total weight exceeds this limit?

```{r}
set.seed(123)

#defining
men_mean <- 5.13 #mean of logarithmic weight of men
men_sd <- .17 #standard deviation of log weight of men
fem_mean <- 4.96 #mean of logarithmic weight of female
fem_sd <- .20 #standard deviation of log weight of female

#number of ppl, capacity, number of simulations
n <- 10
cap <- 1750
num_sims <- 10000	

Above_capacity <- rep(NA, num_sims)	#vector of lenght 'num_sum' filled with NA, will be used to store whether each simulation resulted in a total weight exceeded the capacity

for (x in 1:num_sims) {	
  # runs the simulation "num_sims" times 
  weight_tot <- rep(NA, n)	#this vector stores the info about the weight of each adult
  is_female <- rbinom(n, 1, 0.49)#binary vector representing gender, probability of this being a male is 0.51
  for (i in 1:n) {	
    # inner loop runs for each adult in the simulation generating weight and calculating total
    log_weight <- ifelse(is_female[i]==1,	
    # if the adult is fem then the weight is drawn from fem data
        rnorm(1, fem_mean, fem_sd),	
    
        rnorm(1, men_mean, men_sd)	
    # if the adult is male then the weight is drawn from men data
                         )	
    weight_tot[i]<- exp(log_weight)	
    # value of total weights is calculated as sum of weights of adults and stored in weight_tot vector
  }	
  Above_capacity[x] <- ifelse(cap<sum(weight_tot), 1, 0)	
  # if the total weight exceeds the elevator capacity (set at 1750) then the vector is set to 1, otherwise to 0
}	
answer <-  sum(Above_capacity) / num_sims 

#printing answer
cat("The probability of the weight limit of the elevator being exceeded is ", answer, ".\n")
```

---

2. _(5.6 from ROS)_ __Propagation of uncertainty:__ We use a highly idealized setting to illustrate the use of simulations in combining uncertainties. Suppose a company changes its technology for widget production, and a study estimates the cost savings at \$5 per unit, but with a standard error of \$4. Furthermore, a forecast estimates the size of the market (that is, the number of widgets that will be sold) at 40 000, with a standard error of 10 000. Assuming these two sources of uncertainty are independent, use simulation to estimate the total amount of money saved by the new product (that is, savings per unit, multiplied by size of the market).

```{r}
set.seed(128)

# defining parameters
mean_cost_savings_pu <- 5
sd_cost_savings_pu <- 4
mean_market_size <- 40000
sd_market_size <- 10000
number_of_simulations <- 100000

# simulation cost savings
cost_savings_pu <- rnorm(number_of_simulations, mean_cost_savings_pu, sd_cost_savings_pu)

# simulation market size
market_size <- rnorm(number_of_simulations, mean_market_size, sd_market_size)

# simulation overall cost savings
cost_savings <- cost_savings_pu * market_size

# calculating an estimate of the cost savings
estimated_cost_savings <- mean(cost_savings)

#printing answer
cat("The estimated cost savings are ", estimated_cost_savings, "dollars.\n")
```

---

3. _(5.10 from ROS)_ __Inference for a ratio of parameters:__ A (hypothetical) study compares the costs and effectiveness of two different medical treatments.

    - In the first part of the study, the difference in costs between treatments A and B is estimated at \$600 per patient, with a standard error of \$400, based on a regression with 50 degrees of freedom.
    - In the second part of the study, the difference in effectiveness is estimated at 3.0 (on some relevant measure), with a standard error of 1.0, based on a regression with 100 degrees of freedom.
    - For simplicity, assume that the data from the two parts of the study were collected independently.

    Inference is desired for the incremental cost-effectiveness ratio: the difference between the average costs of the two treatments, divided by the difference between their average effectiveness, a problem discussed further by Heitjan, Moskowitz, and Whang (1999).

    (a) Create 1000 simulation draws of the cost difference and the effectiveness difference, and make a scatterplot of these draws.
    (b) Use simulation to come up with an estimate, 50% interval, and 95% interval for the incremental cost-effectiveness ratio.
    (c) Repeat, changing the standard error on the difference in effectiveness to 2.0.
    
```{r}
# defining parameters
mean_cost_difference <- 600
se_cost_difference <- 400
degrees_of_freedom_cost_difference <- 50
mean_effectiveness_difference <- 3.0
se_effectiveness_difference <- 1.0
degrees_of_freedom_effectiveness_difference <- 100

# (a)
simulation_draws <- 1000
set.seed(135)

cost_difference <- rnorm(simulation_draws, mean_cost_difference, se_cost_difference)

effectiveness_difference <- rnorm(simulation_draws, mean_effectiveness_difference, se_effectiveness_difference)

simulation <- data.frame(cost_difference = cost_difference,
                               effectiveness_difference = effectiveness_difference)

ggplot(simulation, aes(x = cost_difference, y = effectiveness_difference)) +
  geom_point() +
  labs(x = "Cost Difference", y = "Effectiveness Difference")

# (b)
simulation_cost_effectiveness_ratio <- cost_difference /effectiveness_difference

estimated_cost_effectiveness_ratio <- mean(simulation_cost_effectiveness_ratio)
cat("The estimated cost-effectiveness ratio is ", estimated_cost_effectiveness_ratio, ".\n")

interval_50 <- quantile(simulation_cost_effectiveness_ratio, c(0.25, 0.75))
cat("The 50% interval is [", interval_50[1], ", ", interval_50[2], "]\n")

interval_95 <- quantile(simulation_cost_effectiveness_ratio, c(0.025, 0.975))
cat("The 95% interval is [", interval_95[1], ", ", interval_95[2], "]\n")

# (c)
se_effectiveness_difference <- 2.0

effectiveness_difference <- rnorm(simulation_draws, mean_effectiveness_difference, se_effectiveness_difference)

simulation <- data.frame(cost_difference = cost_difference,
                               effectiveness_difference = effectiveness_difference)

ggplot(simulation, aes(x = cost_difference, y = effectiveness_difference)) +
  geom_point() +
  labs(x = "Cost Difference", y = "Effectiveness Difference")

simulation_cost_effectiveness_ratio <- cost_difference /effectiveness_difference

estimated_cost_effectiveness_ratio <- mean(simulation_cost_effectiveness_ratio)
cat("The estimated cost-effectiveness ratio is ", estimated_cost_effectiveness_ratio, ".\n")

interval_50 <- quantile(simulation_cost_effectiveness_ratio, c(0.25, 0.75))
cat("The 50% interval is [", interval_50[1], ", ", interval_50[2], "]\n")

interval_95 <- quantile(simulation_cost_effectiveness_ratio, c(0.025, 0.975))
cat("The 95% interval is [", interval_95[1], ", ", interval_95[2], "]\n")
```

---
